---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-images/"
)
```

# survutils

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/survutils)](http://cran.r-project.org/package=survutils)

An R package for Survival Analysis. This package mainly serves as a wrapper around various survival analysis functions (e.g. `coxph`, `survdiff`) to provide a more inituitive interface. It has the following features (grouped by major topics):

Cox Regression

* `get_cox_res`: Run univariate or multivariate cox regression.
* `plot_cox_res`: Generates a forest plot of the univariate or multivariate cox regression results from `get_cox_res`.

Kaplan Meier Estimates/Curves

* `get_surv_prob`: Calculates the survival probability at specified times from a survival curve. 
* `get_nrisk_tbl`: Provides a number at risk table as [typically seen in publications](https://mcfromnz.wordpress.com/2011/11/06/kaplan-meier-survival-plot-with-at-risk-table/).
* `get_logrank_res`: Runs a log-rank test. 

Other

* `get_c_stat`: Calculate C-statistics. 

# How to Install

To install this package using devtools:

```{r, eval = FALSE}
devtools::install_github("tinyheero/survutils")
```

# Cox Regression

`survutils` provides a nice wrapper function `get_cox_res` that allows you to quickly run an univariate or multivariate cox regression on a set of data. The input data is a data.frame for instance (taking the colon dataset from the `survival` R package as the example):

```{r, message = FALSE}
library("survival")
library("knitr")
library("survutils")
library("dplyr")

head(colon) %>%
	select(age, obstruct, time, status, rx) %>%
	kable
```

The relevant columns are:

* `age` and `obstruct`: These are the features we want to regress on.
* `time`: Time to event.
* `status`: Event status (1 for event; 0 for non-event).
* `rx`: Different treatment groups.

Then to run `get_cox_res`:

```{r}
endpoint <- "time"
endpoint.code <- "status"
 
features <- c("age", "obstruct")
cox.res.df <- get_cox_res(colon, endpoint, endpoint.code, features)
kable(cox.res.df)
```

This runs a multivariate cox regression on the entire set of data. We can plot the results using `plot_cox_res`:

```{r, get_cox_res_example}
plot_cox_res(cox.res.df)
```

This gives us a forest plot with the hazard ratio and confidence evidence for each feature. If we are interested in running cox regression within each treatment group, we can make use of the `group` parameter.

```{r}
group <- "rx"
cox.res.df <- get_cox_res(colon, endpoint, endpoint.code, features, group)
kable(cox.res.df)
```

Notice how the output data.frame now has cox regression results for each treatment group (i.e. Obs, Lev, Lev+5FU). We can also plot these results very easily:

```{r, get_cox_res_group_example}
plot_cox_res(cox.res.df, group = group)
```

This will facet the groups so that we can visualize the cox regression results for each treatment group.
